Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-01 paper/disc20clustering.tex

We say that a component or a ground set is \emph{trivial}, if it contains exactly one node.
We say that the node that belongs to trivial component or a ground set is single.
We say that a~configuration is \emph{component respecting}
if the nodes belonging to the same component are collocated.
Similarly, we define a \emph{ground-set-respecting} configuration.

@@

In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
To show it, clusters of capacity $k\geq 3$ are sufficient.
Note that for $k=2$, a lower bound of $3$ is known \cite{repartition-disc}.

@@

The lower bound requires $k\geq 3$.
It is oblivious to the choice of the reconfiguration cost $\alpha$, and we compare the number of performed node exchanges.

@@

\begin{proof}
	Fix any online algorithm \ALG{}.
	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
	\maciek{The problem with the definition! The first pair does not have the same initial location of its nodes.}
	Initially all nodes are single (they belong to their own trivial ground sets).

	In our construction we distinguish a cluster $S_1$, where we form a ground set of size $k-1$.
	Let $x_0$  be the only single node left on $S_1$ and let $y_0$ be an  arbitrary node that does not reside in $S_1$.
	First,
	we construct a ground set $\{x_0, y_0\}$.
	Since \ALG{} has them split, the adversary produces an external request.
	Since $S_1$ contains a ground set of size $k$, \ALG{} must collocate $\{x_0, y_0\}$ in another cluster.
	To preserve a feasible partition of nodes into clusters,
	 replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
	 \maciek{Not covering all cases. $y_0$ may stay in its place. This sentence should be about $x_0$ replaced by $x_1$ only.}

	Next,
	we merge the component of $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
	\ALG{} must collocate $x_1$ and $C$ in a~cluster other than $S_1$ and
	consequently replace $x_1$ with some other (single) node $x_2$.
	We repeat merging the component of single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
	until there are only two single nodes left that  originate from the same cluster.
	Formally, we denote these two single nodes by $x^*, y^*$, and we have $I(x^*) = I(y^*)$, and~for any other pair of single nodes
	$x'$ and $y'$,
	we have $I(x') \neq I(y')$.
	At this point there are at most $\ell+1$ single nodes left,
	otherwise, there would be more pairs of single nodes that were initially in the same cluster.

	Given this sequence of requests,
	the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
	swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
	Hence,
	OPT pays for $2$ node swaps and
	\ALG{} incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate on exclusion?}.
	Finally, $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
\end{proof}

>>>>

We say that a component or a ground set is \emph{trivial}, if it contains exactly one node.
We say that the node that belongs to trivial component or a ground set is \emph{alone}.
We say that a~configuration is \emph{component respecting}
if the nodes belonging to the same component are collocated.
Similarly, we define a \emph{ground-set-respecting} configuration.

@@

In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
It is oblivious to the choice of the reconfiguration cost $\alpha$, and we compare the number of performed node exchanges.
The lower bound requires $k\geq 3$.
Note that for $k=2$, a lower bound of $3$ is known \cite{repartition-disc}.

@@

First, we construct a ground set of size $k-1$ on one of the clusters.
In any configuration, there exists a node that accompanies it.
We only issue requests between such accompanying (time-varying) node, and any node that was initially collocated with it.
Repeating such requests, almost every node must visit this cluster.
In comparison, \OPT performs one migration.

@@

\begin{proof}
	Fix any online algorithm \ALG{}.
	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
	\maciek{A problem with the definition! The first pair does not have the same initial location of its nodes.}
	Initially all nodes are single (they belong to their own trivial ground sets).

	We distinguish a cluster $S_1$, where we form a ground set of size $k-1$.
	We never dismiss any constructed ground set.
	Each cluster host exactly $k$ nodes, hence in any configuration of \ALG, there exists a single alone node in $S_1$.

	We construct a ground set containing this alone node (and some other node).
	Precisely, let $x_0$  be the alone node from $S_1$ and let $y_0$ be an arbitrary node that does not reside in $S_1$.
	First,
	we construct a ground set $\{x_0, y_0\}$.
	Since \ALG{} has them split, the adversary produces an external request.
	Since $S_1$ contains a ground set of size $k$, \ALG{} must collocate $\{x_0, y_0\}$ in another cluster.
	To preserve a feasible partition of nodes into clusters,
	\ALG must replace $x_0$ with another single node $x_1$.

	We continue to construct ground sets containing the current alone node from $S_1$.
	Precisely, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t.~$I(C) = I(x_1)$.
	To preserve a feasible partition of nodes into clusters,
	\ALG{} must collocate $x_1$ and $C$ in a~cluster other than $S_1$ and
	consequently replace $x_1$ with another alone node $x_2$.

	We end this procedure prematurely (before all nodes visit $S_1$) to keep \OPT's cost low.
	We repeat merging the component of single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
	until there are only two single nodes left that  originate from the same cluster.
	Formally, we denote these two single nodes by $x^*, y^*$, and we have $I(x^*) = I(y^*)$, and~for any other pair of single nodes
	$x'$ and $y'$,
	we have $I(x') \neq I(y')$.
	Note that at this point there are at most $\ell+1$ single nodes left,
	otherwise, there would be more pairs of single nodes that were initially in the same cluster.

	Finally, we bound the cost of \OPT.
	Given this sequence of requests,
	\OPT migrates $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
	swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
	Hence,
	OPT pays for $2$ node swaps and
	\ALG{} incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate on exclusion?}.
	Finally, $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
\end{proof}
